Gazing at a Robot's Head Reflects our Tendency to Anthropomorphise

### 2016-09-09 main.tex
Updated stats

Could the way we look at robots betray our anthropomorphic ascriptions to the
machine? We show in the article that indeed, the duration of gaze fixations on
the head of a humanoid robot reflects one's tendency to anthropomorphise the
robot.  We reach this conclusion by building on an novel experimental protocol
involving cognitive priming: participants are shown a set of stimuli that are
visually identical and yet induce different assumptions regarding the cognitive
capabilities of the robot. By correlating the resulting gaze patterns with
answers to established questionnaires on anthropomorphic attributions, we show
that fixations on the head of the robot reflect how we anthropomorphize the
machine. Besides, by comparing this gazing behaviour to a human baseline, we further
show that people with a generally higher tendency to anthropomorphise exhibit
similar gaze patterns when they look either at humans or robots.

@@

We explore hereafter a different approach to measure anthropomorphism, based on
eye-tracking. By looking at the eyes fixation on the head of the robot, we show
significantly different gazing behaviours in two experimental conditions that
differ in terms of the (projected) robot's human-likeness. This eye-tracking
based metric is behavioural -- hence not prone to post-hoc reconstruction by the
participants -- and \emph{in-the-moment} -- mitigating the impact of a reflective
elaboration on the nature of the human-robot interaction. We believe that these
two improvements over existing techniques provide a solid ground for the study
of the socio-cognitive bonds that establish between humans and robots when they
interact.

@@

\begin{itemize}
    \item people who generally tend to anthropomorphise exhibit gazing
        behaviours that are similar when looking either at humans or robots, and
        reciprocally (\h{3}). If this is verified, this would confirm that our
        gazing behaviour towards agents is impacted by our tendency to
        anthropomorphise.

@@

To compare gazing behaviours across the robot and human conditions, we besides
computed a \emph{gaze distribution difference} over the 10 AOIs ($N=10$):

@@

Participants were asked in the post-questionnaire whether the video presented a
\emph{``robot kind of task''} or a \emph{``human kind of task''} on a 5-point
scale (numerical values ranging from -2 to 2, where -2 refers to robot kind of
task). We found a significant mean difference ($M_{\text{machine-like}}=TBD (SD=TBD)$
\vs $M_{\text{human-like}}=TBD (SD=TDB)$) in this
manipulation check ($F(1,54) = 29.0, p < .001$): the participants who watched
stimuli in the \emph{machine-like priming} condition identified the tasks as more
robot-like, while participants who watched the videos with a \emph{human-like
priming} perceived the tasks as more human-like.

@@

Running an ANOVA against the change of anthropomorphic attributions
\deltaant before and after the experiment shows that \deltaant is greater for the participants who watched the
human-like tasks ($\Delta_{\text{machine-like}}=-0.44 (SD=5.23)$
\vs $\Delta_{\text{human-like}}=3.55 (SD=4.64)$, $F(1,54) = 6.54, p < .05$\fixme{put the exact
p value here}). In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with richer cognitive
expectations regarding the robot report a significantly higher tendency to anthropomorphise robots
\emph{in general} (as the questionnaire on anthropomorphic attributions
explicitly referred to \emph{robots in general}).

@@

\subsection{Impact of the Tendency to Anthropomorphise on Human
Gazing Behaviours}

We found a significant negative correlation between
\deltagaze and
the initial tendency to anthropomorphise \anti{} (Pearson Correlation Coefficient $r(54) = -0.43,
p < .001$, figure~\ref{h2}). The higher the (initial) tendency to anthropomorphise, the smaller
the difference in gazing behaviours between the robot and human conditions. This
supports hypothesis \h{2}.

@@

We consider the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf hands}, {\sf
torso} and {\sf legs} (the left and right arms, hands and legs are grouped
together, other AOIs not belonging to the robot are left out). Based on ANOVA results for the robot in both the machine-like and human-like
cognitive priming conditions, the fixations on the head in the human-like condition
were found to be significantly higher than that in the machine-like condition
($F(1,54) = 4.2, p < .05$\fixme{put the exact p value here}). For all other AOIs groups, the fixations across
machine-like and human-like conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

@@

Our experimental protocol enabled us to examine two additional questions: by
administrating the Godspeed questionnaire on anthropomorphism and mind
attribution first as pre-test, we gained a picture of the initial tendency of
our participants to anthropomorphise robots. Using our \deltagaze measure of the
gazing behaviour similarity, we have evidenced that people with an higher
tendency to anthropomorphise robots, also exhibit a more similar gazing
behaviour when they watch either a human or a robot performing the same task. If
confirmed (in particular with balanced human genders), this is a significant
result. The Godspeed questionnaire assesses how likely we are to
anthropomorphise machines in an abstract way: no concrete robots nor interaction
situations are shown; the participants might react to the sentence ``Please rate
your impression of robots'' in widely different ways depending on their own
constructed images of what robots are. And yet, our results show a significant
correlation with gazing behaviour: the more you tend to anthropomorphise, the
more you look at robots the way you look at humans. This implies that the mere
propensity to attribute physical or mental human traits to robots suffice to
influence a mostly unconscious (yet easy to observe) physical behaviour.  This
link between psychological attributions and a resulting physical behaviour might
be of interest to the Human-Robot Interaction community, especially if we endow
the robot with the ability to measure itself, \emph{during} the interaction, the
time spent by the human on its face. This would for instance allow the robot to
build an initial model of the cognitive expectations held by the human regarding
the robot.

>>>>

Could the way we look at robots betray our anthropomorphic ascriptions to the
machine? We show in the article that indeed, the duration of gaze fixations on
the head of a humanoid robot reflects one's tendency to anthropomorphise the
robot.  We reach this conclusion by building on an novel experimental protocol
involving cognitive priming: participants are shown a set of stimuli that are
visually identical and yet induce different assumptions regarding the cognitive
capabilities of the robot. By correlating the resulting gaze patterns with
answers to established questionnaires on anthropomorphic attributions, we show
that fixations on the head of the robot reflect how we anthropomorphize the
machine. Besides, by comparing this gaze behaviour to a human baseline, we further
show that people with a generally higher tendency to anthropomorphise exhibit
similar gaze patterns when they look either at humans or robots.

@@

We explore hereafter a different approach to measure anthropomorphism, based on
eye-tracking. By looking at the eyes fixation on the head of the robot, we show
significantly different gaze behaviours in two experimental conditions that
differ in terms of the (projected) robot's human-likeness. This eye-tracking
based metric is behavioural -- hence not prone to post-hoc reconstruction by the
participants -- and \emph{in-the-moment} -- mitigating the impact of a reflective
elaboration on the nature of the human-robot interaction. We believe that these
two improvements over existing techniques provide a solid ground for the study
of the socio-cognitive bonds that establish between humans and robots when they
interact.

@@

\begin{itemize}
    \item people who generally tend to anthropomorphise exhibit gazing
        behaviours that are similar when looking either at humans or robots, and
        reciprocally (\h{3}). If this is verified, this would confirm that our
        gaze behaviour towards agents is impacted by our tendency to
        anthropomorphise.

@@

To compare gaze behaviours across the robot and human conditions, we besides
computed a \emph{gaze distribution difference} over the 10 AOIs ($N=10$):

@@

Participants were asked in the post-questionnaire whether the video presented a
\emph{``robot kind of task''} or a \emph{``human kind of task''} on a 5-point
scale (numerical values ranging from -2 to 2, where -2 refers to robot kind of
task). We found a significant mean difference ($M_{\text{machine-like}}=-0.46
(SD=1.21)$
\vs $M_{\text{human-like}}=1.16 (SD=1.00)$) in this
manipulation check ($F(1,55) = 30.73, p < .001$): the participants who watched
stimuli in the \emph{machine-like priming} condition identified the tasks as more
robot-like, while participants who watched the videos with a \emph{human-like
priming} perceived the tasks as more human-like.

@@

Running an ANOVA against the change of anthropomorphic attributions
\deltaant before and after the experiment shows that \deltaant is greater for the participants who watched the
human-like tasks ($\Delta_{\text{machine-like}}=2.97 (SD=9.60)$
\vs $\Delta_{\text{human-like}}=9.94 (SD=9.93)$, $F(1,55) = 7.27, p = .009$. In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with richer cognitive
expectations regarding the robot report a significantly higher tendency to anthropomorphise robots
\emph{in general} (as the questionnaire on anthropomorphic attributions
explicitly referred to \emph{robots in general}).

@@

\subsection{Impact of the Tendency to Anthropomorphise on Human
Gaze Behaviours}

We found a significant negative correlation between
\deltagaze and
the initial tendency to anthropomorphise \anti{} (Pearson Correlation
Coefficient $r(55) = -0.38,
p = .003$, figure~\ref{h2}). The higher the (initial) tendency to anthropomorphise, the smaller
the difference in gaze behaviours between the robot and human conditions. This
supports hypothesis \h{2}.

@@

We consider the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf hands}, {\sf
torso} and {\sf legs} (the left and right arms, hands and legs are grouped
together, other AOIs not belonging to the robot are left out). Based on ANOVA results for the robot in both the machine-like and human-like
cognitive priming conditions, the fixations on the head in the human-like condition
were found to be significantly higher than that in the machine-like condition
($F(1,55) = 4.07, p = .04$\fixme{put the exact p value here}). For all other AOIs groups, the fixations across
machine-like and human-like conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

@@

Our experimental protocol enabled us to examine two additional questions: by
administrating the Godspeed questionnaire on anthropomorphism and mind
attribution first as pre-test, we gained a picture of the initial tendency of
our participants to anthropomorphise robots. Using our \deltagaze measure of the
gaze behaviour similarity, we have evidenced that people with an higher
tendency to anthropomorphise robots, also exhibit a more similar gazing
behaviour when they watch either a human or a robot performing the same task. If
confirmed (in particular with balanced human genders), this is a significant
result. The Godspeed questionnaire assesses how likely we are to
anthropomorphise machines in an abstract way: no concrete robots nor interaction
situations are shown; the participants might react to the sentence ``Please rate
your impression of robots'' in widely different ways depending on their own
constructed images of what robots are. And yet, our results show a significant
correlation with gaze behaviour: the more you tend to anthropomorphise, the
more you look at robots the way you look at humans. This implies that the mere
propensity to attribute physical or mental human traits to robots suffice to
influence a mostly unconscious (yet easy to observe) physical behaviour.  This
link between psychological attributions and a resulting physical behaviour might
be of interest to the Human-Robot Interaction community, especially if we endow
the robot with the ability to measure itself, \emph{during} the interaction, the
time spent by the human on its face. This would for instance allow the robot to
build an initial model of the cognitive expectations held by the human regarding
the robot.
